### 2021-03-26 Chapter-3/Chapter-3.tex
Updates from Overleaf

\chapter{Scalable Side-channel Vulnerability Detection}\label{chapter3}
\section{Problem}

@@

>>>>

\chapter{Fast and Precise Side-channel Vulnerability Detection}\label{chapter3}
\section{Problem}
Side-channel attacks infer sensitive information such as cryptographic keys, personal data from software products unconsciously. As we discuss in \S\ref{chapter1} and \S\ref{chapter2}, patching software are usually easier than adopting new hardware to defend against side-channel attacks.

@@

\section{An illustrative example}

### 2021-03-26 Chapter-4/Chapter-4.tex
Updates from Overleaf

\section{Quantification}

@@

However, prior static approaches cannot precisely capture the amount of leakage. According to the definition of Shannon entropy, the leakage will be
$\frac{1}{2^{64}}*\log_{2}\frac{1}{2^{64}} + \frac{2^{64}-1}{2^{64}}
*\log_{2}\frac{2^{64}-1}{2^{64}} \approx 0$ bits. Max-entropy is defined from the
number of possible observations. Because the program has two
branches, tools based on max-entropy will report the code has a $\log_2{2} = 1$
bit leakage.

\section{Approximate Model Counting}
\label{MCreasons}
%According to
%Definition~\ref{def} introduced in \S\ref{sec:trace-qif},
%the problem of quantifying the amount of leaked information can be reduced to
%the problem of computing the number of items in $K^o$. However, we find that while
%there are various propositional model counters (e.g., \#SAT), they are not
%sufficient scalable for production cryptosystem analysis.
%there is no open source modulo theories counter (\#SMT) available.

>>>>

\section{Background}
Given an event $e$ that occurs with the probability $p(e)$, we receive
\begin{displaymath}
    I = - \log_2p(e)
\end{displaymath}
bits of information by knowing the event $e$ happens according to information theory~\cite{shannon1948mathematical}. 
Considering a char variable $a$
with one byte storage size in a C program, its value ranges from 0 to 255.
Assume $a$ has a uniform distribution. If we observe that
$a$ equals $1$, the probability of this observation is $\frac{1}{256}$. So 
we get $-\log(\frac{1}{256}) = 8$ bits information, which is exactly the size
of a char variable in the C program.

Existing works on information leakage quantification typically use Shannon
entropy~\cite{clark2007static,Wichelmann:2018:MFF:3274694.3274741},
min-entropy~\cite{10.1007/978-3-642-00596-1_21}, and max-entropy~\cite{182946,
Doychev:2017:RAS:3062341.3062388}. In these frameworks, the input sensitive
information $K$ is considered as a random variable.

Let $k$ be one of the possible
value of $K$. The Shannon entropy $H(K)$ is defined as
\begin{displaymath}
    H(K) = - \sum_{k {\in} K}p(k)\log_2(p(k))
\end{displaymath}

Shannon entropy can be used to quantify the initial uncertainty about the
sensitive information. It measures the amount of information in a system.

Min-entropy describes the information leaks for a program with the most likely input. 
For example, min-entropy can be used to describe the
best chance of success in guessing one's password using the
most common password. %, which is defined as
\begin{displaymath}
    \mathit{min\text{-}entropy} = - \log_2(p_{\mathit{max}})
\end{displaymath}

Max-entropy is defined solely on the number of possible observations.
%It is equal to $-\log_2{n}$.
\begin{displaymath}
    \mathit{max\text{-}entropy} = -\log_2{n}
\end{displaymath}
As it is easy to compute, most recent works~\cite{182946,Doychev:2017:RAS:3062341.3062388} use max-entropy as the definition of
the amount of leaked information.

To illustrate how these definitions work, we consider the code
fragment in Figure~\ref{fig:side-channel}. It has two secret-dependent
control-flows, A and B.

\begin{figure}[h!]
\centering
\begin{lstlisting}[xleftmargin=.03\textwidth,xrightmargin=.01\textwidth]
uint8_t key[2], t1, t2;
get_key(key);              // 0 <= key[0], key[1] < 256
t1 = key[0] + key[1];
t2 = key[0] - key[1];
if (t1 < 4){               // leakage site A
    foo();    
}                          
if (t2 > 0){               // leakage site B     
    doo();    
}                          
\end{lstlisting}
\vspace*{-1pt}
    \caption{Side-channel leakage}
    \label{fig:side-channel}
\end{figure}
In this paper we assume an attacker can observe the secret-dependent control-flows in Figure~\ref{fig:side-channel}.
Therefore, an attacker can have two different observations for each leak site
depending on the value of the $\mathit{key}$: $A$ for function \textsf{foo} is executed, 
$\neg A$ for function \textsf{foo} is not executed, $B$ for function \textsf{doo} is
executed, and $\neg B$ for function \textsf{doo} is not executed. Now the
question is how much information can be leaked from the above code if an
attacker knows which branch is executed?

\begin{table}[ht]
    \centering\small\footnotesize
    \caption{The distribution of observation}\label{shtable}
    \vspace*{-0pt}
%    \resizebox{\columnwidth}{!}{
    \begin{tabular}{l|cc|cc}
        \hline

        Observation ($o$)   & $A$ & $\neg A$ & $B$ & $\neg B$ \\ \hline
        Number of Solutions & 65526       & 10        & 32768     & 32768           \\ \hline
        Possibility (p)     & 0.9998      & 0.0002    & 0.5    & 0.5       \\
        \hline
    \end{tabular}
%        }
\end{table}

Assuming $\mathit{key}$ is uniformly distributed, we can calculate the corresponding
possibility by counting the number of possible inputs. Table~\ref{shtable}
describes the probability of each observation. We use the above three types of 
leakage metrics to calculate the amount of leaked information for leak A and leak B.

\vspace{3pt}
\textbf{Min Entropy.}
As $p_{A\mathit{max}} = 0.9998$ and $p_{B\mathit{max}} = 0.5$, 
with the definition, min-entropy equals to
\begin{align*}
    \mathit{min\text{-}entropy_A} & = -\log_2{0.9998} = 0.000\ \mathrm{bits} \\
    \mathit{min\text{-}entropy_B} & = -\log_2{0.5} = 1.000\ \mathrm{bits}
\end{align*}

\textbf{Max Entropy.}
Depending on the value of key, the code can run two different branches for each leakage site. 
Therefore, with the max entropy
definition, both leakage sites leak

\begin{displaymath}
    \mathit{max\text{-}entropy} = -\log_2{2} = 1.000\ \mathrm{bits}
\end{displaymath}

\textbf{Shannon Entropy.}
Based on Shannon entropy, the respective amount of information in A and B equals to
{\footnotesize
\begin{align*}
    \mathit{Shannon\text{-}entropy_A} & = -(0.9998*\log_{2}0.9998      \\
                                    & \qquad+ 0.0002*\log_{2}0.0002)  \\
                                    & = 0.000\ \mathrm{bits}         \\
    \mathit{Shannon\text{-}entropy_B} & = -(0.5*\log_{2}0.5      \\
                                    & \qquad+ 0.5*\log_{2}0.5)        \\
                                    & = 1.000\ \mathrm{bits}                             
\end{align*}
}
In the next section, we will show that these measures work well only
theoretically in a static analysis setting. 
Generally, they do not apply to dynamic analysis or real
settings. We will present that the static or theoretical results could be
dramatically different from the real world, and we do need a better method to
quantify the information leakage from a practical point of view.

\section{Leakage Definition}

@@

However, prior static approaches cannot precisely capture the amount of leakage. According to the definition of Shannon entropy, the leakage will be
$\frac{1}{2^{64}}*\log_{2}\frac{1}{2^{64}} + \frac{2^{64}-1}{2^{64}}
*\log_{2}\frac{2^{64}-1}{2^{64}} \approx 0$ bits. Max-entropy is defined from the
number of possible observations. Because the program has two
branches, tools based on max-entropy will report the code has a $\log_2{2} = 1$
bit leakage.
Both approaches fail to tell how much information is leaked during the execution
precisely. The problem with existing methods is that they are static-based so
input values and real runtime information are neglected by their approaches. 
They assume an attacker runs
the program multiple times with many different or random sensitive inputs. As
shown in Figure~\ref{fig:gap}(a), previous models, both Shannon entropy and max
entropy, give an ``average'' estimate of the information leakage. However, it is
not the typical scenario for an adversary to launch a side-channel attack. When
a side-channel attack happens, the adversary wants to retrieve the sensitive
information, in which case the sensitive information is fixed (e.g., AES keys).
The adversary will run the attack over and over again with fixed input and 
guess the value bit by
bit (e.g., Kocher's timing attacks~\cite{kocher1996timing}), as in Figure~\ref{fig:gap}(b). We want to have a
theory for dynamic analysis that if the theory says an attack leaks $x$ bits of
secret information from a side-channel vulnerability, then $x$ should be useful
in estimating the sensitive level of the vulnerability. However, the above
methods all fail in real attack models. This is the first challenge we face
\textbf{(Challenge C1)}.

\begin{figure}
    \centering
    \includegraphics[width=.65\columnwidth]{./figures/RA.pdf}
\vspace*{-6pt}
    \caption{The gap between the real attack and previous models}\label{fig:gap}
\end{figure}

\subsection{Theoretical Analysis \textbf{(Solution to Challege C1)}}
Now we present our metric to quantify the amount of leaked information from
dynamic analysis.

In information theory, the mutual information (I) is a measure of the mutual
dependence between two variables. Here we use I to describe the
dependence between original sensitive keys ($K$) and attackers' observations ($O$), 
which is defined as:

\begin{equation} \label{eq:1}
    I(K;O) = \sum_{k {\in} K}{\sum_{o {\in} O}{p(k, o)\log_2\frac{p(k, o)}{p(k)p(o)}}}
\end{equation}

where $P(k_i, o_i)$ is the joint discrete distribution of $K$ and $O$.
Alternatively, the mutual information can also be equivalently expressed as:

\begin{equation} \label{eq:2}
    I(K;O) = H(K) - H(K|O)
\end{equation}

$H(K|O)$ is the entropy of $K$ with the condition $O$. It quantifies the
uncertainty of $K$ given the value of $O$. In other word, the conditional
entropy $H(K|O)$ marks the uncertainty about $K$ after an adversary has gained
some observations ($O$).
\begin{equation}
    H(K|O) = - \sum_{o {\in} O} {p(o) \sum_{k {\in} K}{p(k|o)\log_2p(k|o)}}
\end{equation}

In this project, we hope to give a very precise definition of information
leakages. Suppose an attacker runs the target program with one
fixed input, we want to know how much information he can infer by observing the
memory access patterns ($o$). We come to the simple slogan
~\cite{10.1007/978-3-642-00596-1_21} %% where the information
%% leakage equals:
%% \textbf{Initial uncertainty - remaining uncertainty}
that
\begin{align*}
     & \mathit{Information\ leakage} =                                         \\
     & ~~~~ \mathit{Initial\ uncertainty} - \mathit{Remaining\ uncertainty}.
\end{align*}

Next we compare the Eq.~(\ref{eq:2}) with the above slogan, we find $H(K)$
is the $\mathit{Initial\ uncertainty}$ and $H(K|O)$ is $\mathit{Remaining\
uncertainty}$. During a real attack, the observation ($o$) is known.  Thus we
have $H(K|O) = H(K|o)$.

Therefore, we define the amount of leaked information as
\begin{displaymath}
    Leakage = H(K;o) = H(K) - H(K|o)
\end{displaymath}

For a program ($\beta$) without knowing any domain information, all possible sensitive
inputs should appear equally. Therefore, for any $k \in K$, $p(k) =
\frac{1}{|K|}$. So we have
$$H(K) = \sum_{k {\in} K}\frac{1}{|K|}\log_2{|K|} = \log_2{|K|}$$
For any $k' \in K \setminus K^o$, $p(k'|o) = 0$. We get
\begin{align*}
    H(K;o) & = - \sum_{k {\in} K^o}{p(k|o)\log_2p(k|o)}                         \\
           & \qquad   - \sum_{k` {\in} (K \setminus K^o)}{p(k'|o)\log_2p(k'|o)} \\
           & = \sum_{k {\in} K^o}\frac{1}{|K^o|}\log_2{|K^o|}                   \\
           & = \log_2{|K^o|}
\end{align*}

\begin{mydef}
    \label{def}
    Given a program $\beta$ with the input set $K$,
    an adversary has the observation $o$ when the input $k{\in}K^o$.
    We denote it as
    $$\beta(K^o, m) \rightarrow	o$$

    The amount of leaked information $L_{\beta(k)\rightarrow o}$ based on the observation ($o$) is
    $$L_{\beta(k)\rightarrow o} = \log_2{|K|} - \log_2{|K^o|}$$
\end{mydef}

The above definition~\cite{AskarovC12} can be understood in an intuitive way. Suppose an attacker
wants to guess a 128-bit encryption key from a program. 
Without any domain knowledge, 
he can find the key by performing exhaustive search over $2^{128}$ possible keys. 
However, the program has a side-channel leakage site. After the program finishes execution, the
attacker get some leaked information and only need to find the key by performing 
exhaustive search over $2^{120}$ possible keys. Then we can say that 8 bits of the information
is leaked. In this example, $2^{128}$ is the size of $K$ and $2^{120}$ is the size of $K^o$.

With the definition, if an attacker observes that the code in
Figure~\ref{fig:password-checker} runs the branch 1, then the $K^{o^{1}} =
\{\mathrm{``password"}\}$. Therefore, the information leakage $L_{P(k)=o^{1}} =
\log_2{2^{64}} - \log_2{1} = 64$ bits, which means the key is totally leaked. If
the attacker observes the code hits branch 2, the leaked information is
$L_{P(k)=o^{2}} = \log_2{2^{64}} - \log_2{(2^{64}-1)} \approx 0$ bit.

We can also calculate the leaked information from the sample code in
Figure~\ref{fig:side-channel}. As the size of input sensitive information is
usually public. The problem of quantifying the leaked information has been
transferred into the problem of estimating the size of input key $|K^o|$ under
the condition $o \in O$. The result is shown in Table~\ref{shtable2}. We can see
that some branches (e.g., A) or traces leak much more information than some others. 
Those kinds of branches can be vulnerable when an attacker's data is incorporate. 
For example, the code will skip some of the calculation if the value is 0 in big 
number multiplication.

In
contrast, an \emph{average} estimate based on random secret input information is around
1 bit, as shown in \S\ref{sec:background_leak} and Table~\ref{shtable}, is
not very useful in practice as an attacker is able to get much more leaked
information in some attack scenarios.

\begin{table}[ht]
    \centering\small\footnotesize
    \caption{New leakage modeling results}
    \label{shtable2}
    \vspace*{-9pt}
%    \resizebox{\columnwidth}{!}{
    \begin{tabular}{l|cc|cc}
        \hline

        Observation ($o$)   & $A$ & $\neg A$ & $B$ & $\neg B$ \\ \hline
        Number of Solutions & 65526       & 10        & 32768     & 32768           \\ \hline
        Leaked Information (bits) & 0.0         & 14.7      & 1.0       & 1.0         \\
        \hline
    \end{tabular}
\end{table}
\section{Approximate Model Counting}
\label{MCreasons}
%According to
%Definition~\ref{def} introduced in \S\ref{sec:trace-qif},
%the problem of quantifying the amount of leaked information can be reduced to
%the problem of computing the number of items in $K^o$. However, we find that while
%there are various propositional model counters (e.g., \#SAT), they are not
%sufficient scalable for production cryptosystem analysis.
%there is no open source modulo theories counter (\#SMT) available.
